From Freebase to Wikidata: The Great Migration

### 2015-10-14 paper.tex
Rewrote section on backend. Still working on a small architecture diagram.

\begin{figure}[!htbp]
    \centering
    \begin{subfigure}[b]{1.0\columnwidth}
        \includegraphics[width=\textwidth]{img/primary-sources.pdf}
        \caption{Incoming references}
        \label{fig:barack-obama}
    \end{subfigure}
    \begin{subfigure}[b]{1.0\columnwidth}
        \includegraphics[width=\textwidth]{img/wikidata-cause-of-death.pdf}
        \caption{Incoming claims and references}
        \label{fig:cause-of-death}
    \end{subfigure}
    \caption{Primary Sources Tool user interface, incoming data from Freebase shown in light blue}
    \label{fig:primary-sources-tool}
\end{figure}

@@

\subsection{Back-end of the Tool}

\todo{Sebastian or Thomas: write me}

>>>>

\begin{figure}[!htbp]
    \centering
    \begin{subfigure}[b]{1.0\columnwidth}
        \includegraphics[width=\textwidth]{img/primary-sources.pdf}
        \caption{Incoming references}
        \label{fig:barack-obama}
    \end{subfigure}
    \begin{subfigure}[b]{1.0\columnwidth}
        \includegraphics[width=\textwidth]{img/wikidata-cause-of-death.pdf}
        \caption{Incoming claims and references}
        \label{fig:cause-of-death}
    \end{subfigure}
    \caption{Primary Sources Tool user interface. Statements coming from
      Freebase and not yet available in Wikidata are shown in light blue. Users
      can approve or reject these statements using the ``approve''/``reject'' buttons.}
    \label{fig:primary-sources-tool}
\end{figure}

@@

\subsection{Backend of the Tool}

The objective of the backend was to have a REST API allowing us to serve data
efficiently to the front-end of the tool and providing statistical data about
the approvals and rejections done over time. As the resources of the Wikimedia
foundation are limited, the backend has been implemented in C++ and runs as a
fastcgi application in the lighttpd webserver.

Entities and statements are stored in an SQL database (MySQL or SQLite) and
follow the Wikidata data structure with an additional status field to keep track
of statements that have already been approved or rejected, or not yet
visited. Statements can be grouped in ``datasets'' and ``uploads'' to
distinguish data from different sources.\footnote{At the time of this writing,
  the backend contained 5 different datasets}

The REST API of the backend supports the following types of queries:
\begin{itemize}
\item Get an entity (i.e. its statements) by Wikidata QID
  (e.g. \verb|GET /entities/Q2529789|).
\item Get a random entity with unapproved (or approved, rejected) statements
  (\verb|/entities/any|).
\item Get a statement by database ID (\verb|GET /statements/1|) or get a list of
  random statements (\verb|GET /statements/any|).
\item Set a new state for a statement \\(\verb|POST /statements/1?state=approved&user=Alice|).
\end{itemize}
All GET requests support an additional \Verb|state=...| query parameter to
filter by state. This allows Wikidata developers to analyse statements that have
already been updated (e.g. ``return 10 random rejected statements'') for further
analysis. The default of this parameter is \verb|unapproved|.

In addition to these main request types, the REST API also provides endpoints
for importing and deleting datasets, as well as for retrieving system
status.\footnote{The system status is also presented in more convenient form at
  \url{http://tools.wmflabs.org/wikidata-primary-sources/status.html}}

All update activity around statements is logged in an activity log for
statistical purposes using the Wikidata user name of the user who performed the
update. Activity information can be displayed in an aggregated form (top users,
top rejecters, trends) to give to the Wikidata contributors a more gamified
experience.

\todo{Review and expand}
